---
title: "Statistical Inference - Basic Inferential Data Analysis"
author: "H.Harvey"
date: "25 December 2015"
output: word_document
---
```{r setoption, echo=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
```

## Introduction
The purpose of this project is to investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials.

## Setting the variables
```{r}
## R Preparation for reproducible results
set.seed(10)
# set lambda to 0.2
lambda <- 0.2
# 40 samples
Num_Sample <- 40
# 1000 simulations
Num_Sim <- 1000
```

## Calculations of the sample mean
```{r}
# Replicate 1000 instances of the random 40 exponentials
Sim_Exp <- replicate(Num_Sim, rexp(Num_Sample, lambda))
# Calculate the mean of the sample of exponentials
mean_Sim_Exp <- apply(Sim_Exp, 2, mean)
# Calculate the mean of the sample mean
mean_Sample <- mean(mean_Sim_Exp)
# Calculate the theoritical mean
mean_Theory <- 1/lambda
```

## Graphical representation of the central/theoritical mean
```{r}
# Represent the sample mean distribution
hist(mean_Sim_Exp, breaks=50, xlab = "Mean", main = "Sample mean distributions")
# Represent the sample mean of the random simulation
abline(v = mean_Sample , col = "blue")
# Represent the theoritical mean
abline(v = mean_Theory, col = "orange")
legend('topright', c("Simulation mean", "Theoretical mean"),
       lty=c(1,1), col=c("blue", "orange"))

round(mean_Sample, digits = 3)
mean_Theory
```

The distribution of sample means is centered at 5.045 and the theoretical center of the distribution is calculated at 5. The center of distribution of averages of 40 exponentials is very close to the theoretical center of the distribution.

## Calculation of the variance and standard variation
```{r}
# Calculate the standard deviation of the sample distribution
stddev_Sample <- sd(mean_Sim_Exp)
# Calculate the  theoritical standard deviation
stddev_Theory <- (1/lambda)/sqrt(Num_Sample)
# Calcultate the variance of the sample distribution
var_Sample <- stddev_Sample^2
# Calculate the theoritical variation
var_Theory <- ((1/lambda)*(1/sqrt(Num_Sample)))^2
```

```{r}
round(stddev_Sample, digits = 3)
round(stddev_Theory, digits = 3)
round(var_Sample, digits = 3)
var_Theory
```

The standard variation is 0.798 where the theoeritical standard deviation is calculated at 0.791. The variance of sample means is 0.637 where the theoretical variance of the distribution is calculated at 0.625.

## Graphical representation of the distribution
```{r}
hist(mean_Sim_Exp, breaks=50, prob=TRUE,
     main="Density Simulation",
     xlab="")
# Calculate and identify the density of the sample mean
lines(density(mean_Sim_Exp), col="blue")
# Identify the mean of the sample mean
abline(v=mean(mean_Sim_Exp), col="blue")
# Identify the theoretical mean
abline(v=1/lambda, col="orange")

# Calculate the theoretical density of the averages of samples
xfit <- seq(min(mean_Sim_Exp), max(mean_Sim_Exp), length=100)
yfit <- dnorm(xfit, mean=mean_Theory, sd=stddev_Theory)
lines(xfit, yfit, pch=22, col="orange", lty=2)

# Legend
legend('topright', c("Simulation distribution", "Theoretical distribution"),
       lty=c(1,2), col=c("blue", "orange"))
```

As represented in the graph, the mean of the random sampled exponantials distributions overlaps with the theoritical mean of the distribution, due to the Central Limit Theorem.

## Validation
```{r}
qqnorm(mean_Sim_Exp)
qqline(mean_Sim_Exp, col = 2)
```

## Conclusion
Following the central limit theorem (CLT), we can claim that the distribution of averages of 40 exponentials is very close to a normal distribution.
